---
output:
  pdf_document: default
  html_document: default
---
---
title: 'Class6: R Functions'
author: "R(PID:A59010419"
date: "10/15/2021"
output: html_document

# Quick Rmarkdown Tutorial 


We can Write text. **Style** or *style*

student1 <- c(100, 100, 100, 100, 100, 100, 100, 90) 
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)

```{r}
plot(1:10)
```
```{r}
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90) 
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
student3 <- c(90, NA, NA, NA, NA, NA, NA, NA)

```
First I think we should Find if there is an NA and set it to 0 
then find and keep the place in the array that arent the minimum (logical)

```{r}
student<-student1
student[is.na(student)] = 0
a=student[-which.min(student)]
mean(a)

```
The is.na() function returns logical True if NA
my intution is what we ended up doing as a class.
```{r}
studentp<-student2
studentp[is.na(studentp)] = 0
mean(studentp[-which.min(studentp)])
```

```{r}
studentp<-student3
studentp[is.na(studentp)] = 0
mean(studentp[-which.min(studentp)])
```
This works for all the students
lets make the functuin 
actually lets make numeric first
```{r}
x <- student3
x <- as.numeric(x)
x[is.na(x)]=0
mean(x[-which.min(x)])
```

okay now function 
```{r}
# Should we drop the lowest 1 for true 0 for false 
grade <- function(x,drop) {
x <- as.numeric(x)
x[is.na(x)]=0
if (drop ==1){
  mean(x[-which.min(x)])}
  else{
  mean(x)
}
  
}
```
test
```{r}
grade(student1,0)
grade(student2,0)
grade(student3,0)
```
now lets play with this finle
```{r}
gradebook <- "https://tinyurl.com/gradeinput"
scores <- read.csv(gradebook, row.names=1)
scores
```
Cool we have a data set now

We havent used this function yet but APPLY() applies function to our data set... this is cool because it replaces a loop

```{r}
ans <- apply(scores,MARGIN=1,grade)

```
**Q2** top scorer 
```{r}
which.max(ans)
```
**Q3** Which HW was the hardest (lowest score)
```{r}
apply(scores,2,grade)
```
Well this drops one lowest score oops
replace NA in grade book 
```{r}
mask<-scores
mask[is.na(mask)]=0
```
Now function, I think sum is better than average personally

```{r}
sumscore<-apply(mask,2,sum)
which.min(sumscore)
```
Homework 2 was the hardest probably.





okay we can use the mean
```{r}
sumscore<-apply(mask,2,mean)
which.min(sumscore)
```

Q4
```{r}
cor(mask$hw5,ans)
```
Or use apply 
apply(mask,2,cor)
```{r}
corvec<-apply(mask,2,cor,ans)
which.max(corvec)
```






