---
title: "Class17"
author: "R(PID:A59010419"
date: "11/24/2021"
output:
  pdf_document: 
    latex_engine: xelatex
---




Lets read in some COVID-9 Vaccine data 

```{r}
vax <- read.csv( 'covid19vaccinesbyzipcode_test.csv' )
head(vax)
```

Q How many entries do we have?
```{r}
nrow(vax)
```



Q1. What column details the total number of people fully vaccinated? 
persons_fully_vaccinated
Q2. What column details the Zip code tabulation area? 
zip_code_tabulation_area
Q3. What is the earliest date in this dataset? 
2021-01-05
Q4. What is the latest date in this dataset? 
2021-11-23

```{r}
skimr::skim(vax)
```

Q5. How many numeric columns are in this dataset? 
9
Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column? 
8355
Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)? 
10.07
Q8. [Optional]: Why might this data be missing?
```{r}
8355/82908 
```
Notice some of these columns is in the date column:
Lets use lubridate
```{r}
library(lubridate)
today()


```
How many days since the first entry in the dataset?

```{r}
vax$as_of_date[1]
```
```{r}
d <- ymd(vax$as_of_date)
```
```{r}
today()-d[1]
```

Write over data set to have this date format

```{r}
vax$as_of_date <- ymd(vax$as_of_date)


```
```{r}
vax$as_of_date[nrow(vax)]

today() - vax$as_of_date[nrow(vax)]
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```
Last update 
How recently 
How many days does the data set span 



```{r}
length(unique(vax$zip_code_tabulation_area))
```

To work with zip codes we can use zipcoderR

```{r}
library(zipcodeR)
geocode_zip('92037')
zip_distance('92037','92109')
reverse_zipcode(c('92037', "92109") )
```


Lets focus on San Deigo County 
```{r}
library(dplyr)

sd <- filter(vax, county == "San Diego")

nrow(sd)
```

```{r}
sd.10 <- filter(vax, county == "San Diego" &
                age5_plus_population > 10000)
```


```{r}
length(unique(sd$zip_code_tabulation_area))
```

Q11. How many distinct zip codes are listed for San Diego County? 
107

```{r}
sd$zip_code_tabulation_area[which.max(sd$age12_plus_population)]

```


Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset? 
Hint
92154


Q What is the average vaccination rate of San Diego County
```{r}
sd.now <- filter(vax, county == "San Diego", as_of_date == "2021-11-23")
head(sd.now)
summary(sd.now$percent_of_population_fully_vaccinated)
```


```{r}
hist(sd.now$percent_of_population_fully_vaccinated)
```



```{r}
ucsd <- filter(sd.now, zip_code_tabulation_area == "92037")
ucsd[1,]$age5_plus_population
```
Q What is the population of the 92037 zip code area and what is the average vaccination 
```{r}
ucsd$percent_of_population_fully_vaccinated
```

```{r}
pb <- filter(sd.now, zip_code_tabulation_area == "92109")
pb$percent_of_population_fully_vaccinated
```

```{r}
ucsd2 <- filter(vax, zip_code_tabulation_area == "92037")
```


```{r}
library(ggplot2)
ggplot(ucsd2) +
  aes(ymd(ucsd2$as_of_date),
      ucsd2$percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x="date", y="Percent Vaccinated")
```

```{r}
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2021-11-16")
summary(vax.36$percent_of_population_fully_vaccinated)
```
Mean .6640 

```{r}
hist(vax.36$percent_of_population_fully_vaccinated)
```


```{r}
sd.36 <- filter(sd, age5_plus_population > 36144 &
                as_of_date == "2021-11-16")
summary(sd.36$percent_of_population_fully_vaccinated)
```


```{r}
vax.36.all <- filter(vax, age5_plus_population > 36144 )
length(unique(vax.36.all$zip_code_tabulation_area))

ggplot(vax.36.all) +
  aes(ymd(vax.36.all$as_of_date),
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2) +
  ylim(c(0,1.0)) +
  labs(x="Date", y="Precent Vaccinated",
       title="Vaccination rate across California",
       subtitle="populations > 36144") + 
      geom_hline(yintercept = 0.67, color="red")


```





